Analysis of 2005-2015 Earthquake Sequence in Northern Iran Using Visibility Graph Approach

### 2016-08-09 abstract.tex
commit before adding methodology.

\begin{abstract}
The seismicity of northern Iran between 2005 and 2015 is investigated by means of the visibility graph (VG) method. For purposes of this study, northern Iran is divided into three tectonic seismic regions: Azerbaijan, Alborz, and Kopeh Dagh. Using the declustered catalog, we studied the VG properties of the magnitude time series. The results show a relationship between the $k-M$ slope and the $b-value$ of the Gutenberg-Richter law. Topological properties (i.e.  $< T_c >$, $k-M$ slope) of the network and dynamic properties of magnitude time series (i.e. $b-value$) significantly decrease before large earthquakes. Combining the results of this study with two similar previous studies improves the linear correlation coefficient factor between $k-M$ and $b-value$, and empowers the idea of a universal relationship between $b-value$ and $k-M$ slope. The behaviors of the VG's properties are similar to previous studies and could be considered an alternative method for analyzing earthquake sequences. 
\end{abstract}

>>>>

\begin{abstract}
The seismicity of northern Iran between 2005 and 2015 is investigated by means of the visibility graph (VG) method. For purposes of this study, northern Iran is divided into three tectonic seismic regions: Azerbaijan, Alborz, and Kopeh Dagh. Using the declustered catalog, we studied the VG properties of the magnitude time series. The results show a relationship between the $k-M$ slope and the $b-value$ of the Gutenberg-Richter law. Topological properties (i.e.  $< T_c >$, $k-M$ slope) of the network and dynamic properties of magnitude time series (i.e. $b-value$) significantly decrease before large earthquakes. Combining the results of this study with three similar previous studies improves the linear correlation coefficient factor between $k-M$ slope and $b-value$, and empowers the idea of a universal relationship between $b-value$ and $k-M$ slope. The behaviors of the VG's properties are similar to previous studies and could be considered an alternative method for analyzing earthquake sequences. 
\end{abstract}

### 2016-08-09 acknowledgements.tex
commit before adding methodology.

\section{Acknowledgements}
\noindent
We greatly appreciate the International Institute of Earthquake Engineering and Seismology (IIEES) in Tehran for maintaining and providing access to the seismic catalog. This research was possible thanks to support by the Center for Earthquake Research and Information (CERI) at the University of Memphis. CERI is designated as a Center of Excellence by the Tennessee Board of Regents and is funded in part by the State of Tennessee under State Sunsent Laws (SB 1510 and HB 1608, 2015{2016).

>>>>

\section{Acknowledgements}
\noindent
We greatly appreciate the International Institute of Earthquake Engineering and Seismology (IIEES) in Tehran for maintaining and providing access to the seismic catalog. This research was possible thanks to support by the Center for Earthquake Research and Information (CERI) at the University of Memphis. CERI is designated as a Center of Excellence by the Tennessee Board of Regents and is funded in part by the State of Tennessee under State Sunsent Laws (SB 1510 and HB 1608, 2015--2016).

### 2016-08-09 introduction.tex
commit before adding methodology.

In this study, we use visibility graph analysis to investigate the earthquake sequence of northern Iran. We divided northern Iran into three tectonic seismic regions, namely Azerbaijan (Az), Alborz Mountain Range (Al), and Kopeh Dagh (Ko).  Recently, northern Iran has been studied in detail from different seismological points of view.  \citet{Nemati2015} studied the most recent 200 years of seismicity in northern Iran. The frequency of shocks varied widely, from one main shock per 6 years (0.17 event/ year) for the Azerbaijan region to 13 earthquakes per 4 years (3.25 event/year) for the Kopeh-Dagh region \citep{Nemati2015}. Using the smoothed seismicity method, \citet{Khoshnevis2016} studied the seismic hazard of the three northern tectonic seismic regions. Having a considerable number of recorded earthquakes, extensive seismic history, and being in a unique tectonic seismic region, northern Iran is a good platform from which to analyze the visibility graph method's capabilities for studying the magnitude time series. Significant aspect of our work include the compassion of earthquake catalog, conversion of the earthquake magnitude and computation of the completeness magnitude for each tectonic seismic zones. We present the relationship between  $k-M$  slope and $b-value$  in these three tectonic seismic regions. We analyze the sensitivity of the catalogs to the number of events and the threshold magnitude. We also investigate the variation of  $<T_C>$  value through time.

>>>>

In this study, we use visibility graph analysis to investigate the earthquake sequence of northern Iran. We divided northern Iran into three tectonic seismic regions, namely Azerbaijan (Az), Alborz Mountain Range (Al), and Kopeh Dagh (Ko).  Recently, northern Iran has been studied in detail from different seismological points of view.  \citet{Nemati2015} studied the most recent 200 years of seismicity in northern Iran. The frequency of shocks varied widely, from one main shock per 6 years (0.17 event/ year) for the Azerbaijan region to 13 earthquakes per 4 years (3.25 event/year) for the Kopeh-Dagh region \citep{Nemati2015}. Using the smoothed seismicity method, \citet{Khoshnevis2016} studied the seismic hazard of the three northern tectonic seismic regions. Having a considerable number of recorded earthquakes, extensive seismic history, and being in a unique tectonic seismic region, northern Iran is a good platform from which to analyze the visibility graph method's capabilities for studying the magnitude time series. Significant aspect of our work includes the compassion of earthquake catalog, conversion of the earthquake magnitude and computation of the completeness magnitude for each tectonic seismic zone. We present the relationship between  $k-M$  slope and $b-value$  in these three tectonic seismic regions. We analyze the sensitivity of the catalogs to the number of events and the threshold magnitude. We also investigate the variation of  $<T_C>$  value through time.

### 2016-08-09 results.tex
commit before adding methodology.

\begin{table}[h]
\centering
\caption{Declustered number of earthquakes, seismic parameters and $k-M$ slope values for north Iran tectonic seismic zones.}
\begin{tabular}{ccccc}
Region          & Number of Earthquake &  $Mc$ &  $b-value$ & $k-M$ slope \\ \hline
Azerbaijan     & 93                                 & 3.5   & 1.0190  & 9.2004       \\ \hline
Alborz            & 794                               & 2.6   & 0.7717  & 9.0461      \\ \hline
Kopek Dagh  & 282                               & 2.6   & 0.5799  & 6.5818     \\ \hline
\end{tabular}
\label{tab:b_k_m_param}
\end{table}

@@

\begin{equation}
b = \frac{log_{10}(e) }{\overline{M} - M_{min}},
\end{equation}
 \noindent
where $\overline{M}$ is the average magnitude, and  $M_ {min}$ is the minimum magnitude in the sample, which in this case is the completeness magnitude for each tectonic seismic region. The number of sequences and threshold magnitudes is considerably different for all tectonic seismic regions.  \citet{Telesca2013}  studied the effect of the number of earthquakes in the catalog on the  $k-M$  slope. Comparing the result with the Guerrero region with reduced random sequence,  \citet{Telesca2013}  found that the number of sequences did not affect the  $k-M$  slope.  According to  \citet{Telesca2012}, the threshold magnitude has minor effect in the VG parameters. In order to analyze the sensitivity of the catalogs to the number of events and the threshold magnitude, we randomly picked 200 sequences from the catalogs with various sequence sizes. The minimum size of random windows is 150 events. (Note that the number of events in each catalog, before removing the events with magnitude less than the completeness magnitude, is 271, 1262, and 399 for Azerbaijan, Alborz, and Kopeh Dagh regions, respectively.). The maximum size is the size of the entire catalog. For each randomly picked magnitude time series,  we estimate $Mc$; then the $k-M$  and $b-value$ are calculated. Fig.~\ref{fig:random} shows the relationship between  $k-M$  slope and  $b-value$  of the randomly selected data and the statistical parameters for variables. Although changing the number of events and the threshold magnitude is a slight change, the results are fairly well clustered for each tectonic seismic region.

 \begin{figure} [ht]
\centering
\includegraphics[scale=0.5]{figures/pdf/Figure07.pdf} 
\caption{ Relationship between $k-M$ and $b-value$ of three Iranian tectonic seismic regions for 200 random sequence. The numbers on the mean and standard deviation plots are coefficient of variations in percent$(standard \ deviation / mean)$.}
\label{fig:random}
\end{figure}

\noindent
The coefficient of variation of parameters is shown in Fig.~\ref{fig:random}. The  $k-M$  slope shows the higher coefficient of variation which indicates the fact that the  $k-M$  slope is more dependent on the window size and threshold magnitude than on the a-  and  b-values , which suggests that  the $k-M$  slope can better represent the dynamic characteristics of the magnitude-time series. 
 \noindent
 Fig.~\ref{fig:regression} shows the relationship between the  $k-M$  slope and $b-value$ for the three tectonic seismic regions in northern Iran and three other studies of Mexican zones  \citep{Telesca2013}, Pannonia zones  \citep{Telesca2014}, and synthetic data \citet{Telesca2014-pone}. Adding the results of the current study into the results of three previous similar studies improves the regression factor, and empowers the concept of the universal character of the relationship between the  $b-value$  and the  $k-M$  slope, as  \citet{Telesca2014} concluded.

\begin{figure} [ht]
\centering
\includegraphics[scale=0.6]{figures/pdf/Figure08.pdf} 
\caption{ Relationship between $k-M$ slope and $b-value$ of three Iranian tectonic seismic zones (this study) and three other studies of Mexican zones \citep{Telesca2013}, Pannonia zones \citep{Telesca2014}, and synthetic data \citep{Telesca2014-pone}. The lines represent the linear regression fit of data. The correlation coefficients (R) are represented for different combination of data. }
\label{fig:regression}
\end{figure}

>>>>

\begin{table}[h]
\centering
\caption{Declustered number of earthquakes, seismic parameters and $k-M$ slope values for north Iran tectonic seismic zones.}
\begin{tabular}{ccccc}
Region          & Number of Earthquake &  $Mc$ &  $b-value$ & $k-M$ slope \\ \hline
Azerbaijan     & 93                                 & 3.5   & 1.0190  & 9.2004       \\ \hline
Alborz            & 794                               & 2.6   & 0.7717  & 9.0461      \\ \hline
Kopeh Dagh  & 282                               & 2.6   & 0.5799  & 6.5818     \\ \hline
\end{tabular}
\label{tab:b_k_m_param}
\end{table}

@@

\begin{equation}
b = \frac{log_{10}(e) }{\overline{M} - M_{min}},
\end{equation}
 \noindent
where $\overline{M}$ is the average magnitude, and  $M_ {min}$ is the minimum magnitude in the sample, which in this case is the completeness magnitude for each tectonic seismic region. The number of sequences and threshold magnitudes is considerably different for all tectonic seismic regions.  \citet{Telesca2013}  studied the effect of the number of earthquakes in the catalog on the  $k-M$  slope. Comparing the result with the Guerrero region with reduced random sequence,  \citet{Telesca2013}  found that the number of sequences did not affect the  $k-M$  slope.  According to  \citet{Telesca2012}, the threshold magnitude has minor effect in the VG parameters. In order to analyze the sensitivity of the catalogs to the number of events and the threshold magnitude, we randomly picked 200 sequences from the catalogs with various sequence sizes. The minimum size of random windows is 150 events. (Note that the number of events in each catalog, before removing the events with magnitude less than the completeness magnitude, is 271, 1262, and 399 for Azerbaijan, Alborz, and Kopeh Dagh regions, respectively.). The maximum size is the size of the entire catalog. For each randomly picked magnitude time series,  we estimate $Mc$; then the $k-M$  and $b-value$ are calculated. Fig.~\ref{fig:random} shows the relationship between  $k-M$  slope and  $b-value$  of the randomly selected data and the statistical parameters for variables. Although changing the number of events and the threshold magnitude is slightly changing the results, however,  the results are fairly well clustered for each tectonic seismic region.

 \begin{figure} [ht]
\centering
\includegraphics[scale=0.5]{figures/pdf/Figure07.pdf} 
\caption{ Relationship between $k-M$ and $b-value$ of three Iranian tectonic seismic regions for 200 random sequence. The numbers on the mean and standard deviation plots are coefficient of variations in percent $(standard \ deviation / mean)$.}
\label{fig:random}
\end{figure}

\noindent
The coefficient of variation of parameters is shown in Fig.~\ref{fig:random}. The  $k-M$  slope shows the higher coefficient of variation which indicates the fact that the  $k-M$  slope is more dependent on the window size and threshold magnitude than on the a-  and  b-values , which suggests that  the $k-M$  slope can better represent the dynamic characteristics of the magnitude-time series. 
 \noindent
 Fig.~\ref{fig:regression} shows the relationship between the  $k-M$  slope and $b-value$ for the three tectonic seismic regions in northern Iran and three other studies of Mexican zones  \citep{Telesca2013}, Pannonia zones  \citep{Telesca2014}, and synthetic data \citep{Telesca2014-pone}. Adding the results of the current study into the results of three previous similar studies improves the regression factor, and empowers the concept of the universal character of the relationship between the  $b-value$  and the  $k-M$  slope, as  \citet{Telesca2014} concluded.

\begin{figure} [ht]
\centering
\includegraphics[scale=0.6]{figures/pdf/Figure08.pdf} 
\caption{ Relationship between $k-M$ slope and $b-value$ of three Iranian tectonic seismic zones (this study) and three other studies of Mexican zones \citep{Telesca2013}, Pannonia zones \citep{Telesca2014}, and synthetic data \citep{Telesca2014-pone}. The lines represent the linear regression fit of data. The correlation coefficients (R) are represented for different combination of data.}
\label{fig:regression}
\end{figure}

### 2016-08-09 seismicity.tex
commit before adding methodology.

\section{Tectonic seismic regions and seismicity parameters}
\noindent
The Iranian plateau is located on the Himalayan-Alpine seismic belt. It has a long history of large magnitude ($M>7$) earthquakes that are well documented dating back to the eighth century. Based on seismicity parameters and the geologic provinces of the plateau, Iranian earthquakes have been categorized into various seismic zones, ranging from the definition of four to nine major seismic zones in the more traditional studies \citep[e.g.,][]{Stocklin1968,Takin1972,Berberian1976}, and up to twenty to twenty-three tectonic seismic regions in the most elaborate ones \citep[e.g.,][]{Nowroozi1976,Tavakoli1999}. \citet{Mirzaei1998},  dividing Iran into five tectonic regions, called Azerbaijan-Alborz, Kopeh-Dagh, Zagros, Central-East Iran, and Makran. Considering the devision of  \citet{Mirzaei1998} as a reference,  \citet{Karimiparidari2013}  divided the Azerbaijan-Alborz tectonic seismic region into two regions, which are Azerbaijan and the Alborz Mountain Range (hereinafter, Alborz). Azerbaijan, Alborz, and Kopeh-Dagh tectonic seismic regions encompass most of northern Iran.  Fig.~\ref{fig:study_region} shows the study area and tectonic seismic regions. \\

\begin{figure} [ht]
\centering
\includegraphics[scale=1]{figures/pdf/Figure01.pdf} 
\caption{Region of interest and tectonic seismic regions. At the top, the map of Iran and surrounding countries. The study area is located in the gray box. At the bottom, the study area containing seismotectonic provinces after \citet{Mirzaei1998} is presented. Dashed line is the subdivision that is proposed by \citet{Karimiparidari2013}}
\label{fig:study_region}
\end{figure}

@@

\subsection{Magnitude Conversion}
\noindent
The catalog of recorded earthquakes from 2005-15, downloaded from IIEES, reported earthquakes based on different magnitude scales \citep{IIEES}. The  $M_L$ ,  $M_S$, and  $mb$  magnitudes were converted through conversion relationships, as defined in  \citet{Zare2014}. Some data were also recorded in  $M_D$  (duration magnitude). These data are reported by the International Seismological Centre (ISC).   \citet{Deniz2010}  developed a set of empirical equations to convert earthquake magnitudes from  $mb$,  $M_D$,  $M_L$, and  $M_S$  scales to  $M_W$  scale using the orthogonal regression procedure. They used data from different data centers, including the ISC, of earthquakes that occurred in Turkey. In this study, we use the conversion equation of  \citet{Deniz2010}  to convert  $M_D$  to  $M_W$.

@@

\begin{figure} [ht]
\centering
\includegraphics[scale=1]{figures/pdf/Figure03.pdf} 
\caption{Declustered instrumental seismicity map (2005 - 2015) of Northern Iran. Different colors indicate the seismicity of different regions. Size of symbols are proportional to the magnitude of the events.}
\label{fig:seismicity}
\end{figure}

\subsection{Catalog Completeness}
\noindent
Catalog completeness is an important factor in studying an earthquake sequence. This characteristic measured as the minimum magnitude of complete recording for earthquake catalog . The completeness magnitude $(M_c)$ is often determined using simple numerical analysis. Two common approaches are maximum curvature method and (MAXC) and goodness-of-fit test (GFT)(see \citet{Wiemer2000}). We use the Goodness-of-fit test and compute $M_c$ for each region's catalog. According to the GFT method, a complete catalog should follow the Gutenberg-Richter power law distribution of magnitude. The steps for selecting the completeness magnitude for each catalog include: 1) calculating the a-  and  b-  value based on minimum magnitude; 2) generating  synthetic events based on the achieved value when the cumulative number of events obey the power law distribution; and 3) calculating goodness of fit for predicted and observed cumulative numbers for each magnitude bin  \citep{Wiemer2000}. We increase the minimum magnitude and repeate the process in order to calculate the goodness of fit for each magnitude.  \citet{Wiemer2000} assumed a goodness of fit of 90\% as a threshold to select the completeness of the catalog. However,  not all frequency-magnitude distributions reach the 90\% mark. Fig.~\ref{fig:completeness} shows the goodness-of-fit values for three tectonic seismic regions. For each tectonic seismic region, we select the magnitude corresponding to the highest goodness of fit or the first magnitude with the goodness of fit greater than 90\%.

@@

\begin{figure} [ht]
\centering
\includegraphics[scale=0.8]{figures/pdf/Figure05.pdf} 
\caption{Variation of moment magnitude ($M_w$) of 2005-2015 earthquake for north Iran as a function of time. The figure only represents the reclustered events greater than minimum magnitude. The 2012 $M_w 6.4$ East Azerbaijan, 2010 $M_w 5.8$ Damghan, and 2012 $M_w 5.4$ Neyshabur earthquakes are clearly visible as maximum values in Azerbaijan (Az),  Alborz (Az), and Kopeh Dagh (Ko) magnitude-time series.}
\label{fig:mag-time}
\end{figure}

>>>>

\section{Tectonic seismic regions and seismicity parameters}
\noindent
The Iranian plateau is located on the Himalayan-Alpine seismic belt. It has a long history of large magnitude ($M>7$) earthquakes that are well documented dating back to the eighth century. Based on seismicity parameters and the geologic provinces of the plateau, Iranian earthquakes have been categorized into various seismic zones, ranging from the definition of four to nine major seismic zones in the more traditional studies \citep[e.g.,][]{Stocklin1968,Takin1972,Berberian1976}, and up to twenty to twenty-three tectonic seismic regions in the most elaborate ones \citep[e.g.,][]{Nowroozi1976,Tavakoli1999}. \citet{Mirzaei1998},  dividing Iran into five tectonic regions, called Azerbaijan-Alborz, Kopeh-Dagh, Zagros, Central-East Iran, and Makran. Considering the division of  \citet{Mirzaei1998} as a reference,  \citet{Karimiparidari2013}  divided the Azerbaijan-Alborz tectonic seismic region into two regions, which are Azerbaijan and the Alborz Mountain Range (hereinafter, Alborz). Azerbaijan, Alborz, and Kopeh-Dagh tectonic seismic regions encompass most of northern Iran.  Fig.~\ref{fig:study_region} shows the study area and tectonic seismic regions. \\

\begin{figure} [ht]
\centering
\includegraphics[scale=1]{figures/pdf/Figure01.pdf} 
\caption{Region of interest and tectonic seismic regions. At the top, the map of Iran and surrounding countries is presented. The study area is located in the gray box. At the bottom, the study area containing seismotectonic provinces after \citet{Mirzaei1998} is presented. Dashed line is the subdivision that is proposed by \citet{Karimiparidari2013}}
\label{fig:study_region}
\end{figure}

@@

\subsection{Magnitude Conversion}
\noindent
The catalog of recorded earthquakes from 2005-15, downloaded from IIEES, reported earthquakes based on different magnitude scales \citep{IIEES}. The  $M_L$,  $M_S$, and  $mb$  magnitudes were converted through conversion relationships, as defined in  \citet{Zare2014}. Some data were also recorded in  $M_D$  (duration magnitude). These data are reported by the International Seismological Centre (ISC).   \citet{Deniz2010}  developed a set of empirical equations to convert earthquake magnitudes from  $mb$,  $M_D$,  $M_L$, and  $M_S$  scales to  $M_W$  scale using the orthogonal regression procedure. They used data from different data centers, including the ISC, of earthquakes that occurred in Turkey. In this study, we use the conversion equation of  \citet{Deniz2010}  to convert  $M_D$  to  $M_W$.

@@

\begin{figure} [ht]
\centering
\includegraphics[scale=1]{figures/pdf/Figure03.pdf} 
\caption{Declustered instrumental seismicity map (2005 - 2015) of northern Iran. Different colors indicate the seismicity of different regions. Size of symbols are proportional to the magnitude of the events.}
\label{fig:seismicity}
\end{figure}

\subsection{Catalog Completeness}
\noindent
Catalog completeness is an important factor in studying an earthquake sequence. This characteristic measured as the minimum magnitude of complete recording for earthquake catalog. The completeness magnitude $(M_c)$ is often determined using simple numerical analysis. Two common approaches are maximum curvature method and (MAXC) and goodness-of-fit test (GFT)(see \citet{Wiemer2000}). We use the Goodness-of-fit test and compute $M_c$ for each region's catalog. According to the GFT method, a complete catalog should follow the Gutenberg-Richter power law distribution of magnitude. The steps for selecting the completeness magnitude for each catalog include: 1) calculating the a-  and  b-  value based on minimum magnitude; 2) generating  synthetic events based on the achieved value when the cumulative number of events obey the power law distribution; and 3) calculating goodness of fit for predicted and observed cumulative numbers for each magnitude bin  \citep{Wiemer2000}. We increase the minimum magnitude and repeat the process in order to calculate the goodness of fit for each magnitude.  \citet{Wiemer2000} assumed a goodness of fit of 90\% as a threshold to select the completeness of the catalog. However,  not all frequency-magnitude distributions reach the 90\% mark. Fig.~\ref{fig:completeness} shows the goodness-of-fit values for three tectonic seismic regions. For each tectonic seismic region, we select the magnitude corresponding to the highest goodness of fit or the first magnitude with the goodness of fit greater than 90\%.

@@

\begin{figure} [ht]
\centering
\includegraphics[scale=0.8]{figures/pdf/Figure05.pdf} 
\caption{Variation of moment magnitude ($M_w$) of 2005-2015 earthquake for north Iran as a function of time. The figure only represents the declustered events greater than minimum magnitude. The 2012 $M_w 6.4$ East Azerbaijan, 2010 $M_w 5.8$ Damghan, and 2012 $M_w 5.4$ Neyshabur earthquakes are clearly visible as maximum values in Azerbaijan (Az),  Alborz (Az), and Kopeh Dagh (Ko) magnitude-time series.}
\label{fig:mag-time}
\end{figure}
